`r glue::glue('Doublet benchmarking results: {params$dataset}')`

### 2024-05-30 analyses/doublet-detection/template-notebooks/02_explore-benchmark-results.Rmd
fix table function usage to avoid duplication, and consolidate pca function

```{r functions}
print_doublet_table <- function(df, pred_col, is_ground_truth = FALSE) {
  # Function to print a table of doublet counts and a sentence stating the percentage of doublets
  # df is expected to contain `pred_col`, should should be provided as a string
  dub_table <- table(df[[pred_col]])
  dub_percent <- round(dub_table[["doublet"]]/sum(dub_table), 4) * 100

  print(dub_table)

@@

  return(output)
}

plot_pca_d <- function(df, color_column, color_lab) {
  # Plot PC with a discrete color variable, showing doublets on top
  # df is expected to contain columns PC1, PC2, and `color_column`
  # The `color_column` argument should _not_ be provided as a string
  ggplot(doublet_df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(alpha = 0.5) + 
  scale_color_manual(values = c("black", "lightblue")) + 
  # ensure doublets are on top for visibility
  geom_point(
    data = dplyr::filter(doublet_df, {{color_column}} == "doublet"), 
    color = "black"
  ) + 
  labs(color = color_lab) + 
  theme(
    legend.title.position = "top",
    legend.position = "bottom"
  )
}

plot_pca_c <- function(df, color_column, pred_column, color_lab) {
  # Plot PC with a continuous color variable, showing doublets on top
  # df is expected to contain columns PC1, PC2, `color_column`, and `pred_column`
  # The `color_column` and `pred_column` arguments should _not_ be provided as strings
  ggplot(doublet_df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(alpha = 0.5) + 
  scale_color_viridis_c(direction = -1) + 
  # ensure doublets are on top for visibility
  geom_point(
    data = dplyr::filter(doublet_df, {{pred_column}} == "doublet"), 
    aes(color = {{color_column}})
  ) + 
  labs(color = color_lab) + 
  theme(
    legend.title.position = "top",
    legend.position = "bottom"
  )
}

@@

```{r}
percent <- print_doublet_table(doublet_df, "ground_truth", is_ground_truth = TRUE)
```

```{r results = "asis"}
percent
```

```{r}
plot_pca_d(doublet_df, ground_truth, "Ground truth doublets")
```

@@

```{r}
percent <- print_doublet_table(doublet_df, "scdbl_prediction")
```

```{r results = "asis"}
percent
```

@@

```{r fig.width = 8}
d_plot <- plot_pca_d(
  doublet_df, 
  scdbl_prediction,
  "scDblFinder prediction"
)

c_plot <- plot_pca_c(
  doublet_df, 
  scdbl_score,
  scdbl_prediction,
  "scDblFinder score"
)

@@

```{r}
percent <- print_doublet_table(doublet_df, "scrublet_prediction")
```

```{r results = "asis"}
percent
```

@@

```{r fig.width = 8}
d_plot <- plot_pca_d(
  doublet_df, 
  scrublet_prediction,
  "scrublet prediction"
)

c_plot <- plot_pca_c(
  doublet_df, 
  scrublet_score,
  scrublet_prediction,
  "scrublet score"
)

@@

      plot_pca_d(
        doublet_df, 
        !!as.symbol(col_name),
        glue::glue("cxds prediction at {threshold}")
      ) +
      labs(title = glue::glue("cxds prediction at {threshold}"))
    }
)
wrap_plots(prediction_pcas)
```

```{r}
# just need a single score pca for this case
# we'll provide the least stringent predictions; this only affects point layering in the plot
pca_c <- plot_pca_c(
  doublet_df, 
  cxds_score, 
  cxds_prediction_0.5,
  glue::glue("cxds scores")
)
wrap_plots(pca_c)
```

@@

```{r}
percent <- print_doublet_table(doublet_df, "cxds_prediction_0.5")
```

```{r results = "asis"}
percent
```

@@

```{r}
percent <- print_doublet_table(doublet_df, "cxds_prediction_0.75")
```

```{r results = "asis"}
percent
```

@@

```{r}
percent <- print_doublet_table(doublet_df, "cxds_prediction_0.9")
```

```{r results = "asis"}
percent
```

>>>>

```{r functions}
print_doublet_table <- function(df, pred_col, is_ground_truth = FALSE) {
  # Function to print a table of doublet counts and a sentence stating the percentage of doublets
  # df is expected to contain `pred_col`, should should be provided as a string
  dub_tab <- table(df[[pred_col]])
  dub_percent <- round(dub_tab[["doublet"]]/sum(dub_tab), 4) * 100

@@

  return(
    list(
      "table" = dub_tab, 
      "percent" = output
    )
  )
}

plot_pca <- function(df, 
                     color_column, 
                     pred_column = NULL,
                     color_type = "d",
                     color_lab) {
  # Plot PCs colored by either discrete or continuous color variable, showing doublets on top
  # df is expected to contain columns PC1, PC2, `color_column`, and if color_type == "c" also `pred_column`. These should _not_ be provided as strings
  # By default, this function assumes a discrete color variable (`color_type = "d"`) with two levels. Use `color_type = "c"` for continuous.

  p <- ggplot(doublet_df) + 
    aes(x = PC1, 
        y = PC2, 
        color = {{color_column}}) +
  geom_point(alpha = 0.5) + 
  labs(color = color_lab) + 
  theme(
    legend.title.position = "top",
    legend.position = "bottom"
  )

  # Set the palette, and ensure doublets are on top for visibility
  if (color_type == "d") {
    p <- p + 
      scale_color_manual(values = c("black", "lightblue")) + 
    geom_point(
      data = dplyr::filter(doublet_df, {{color_column}} == "doublet"), 
      color = "black"
    )
  } else if (color_type == "c") {
    p <- p + 
      scale_color_viridis_c(direction = -1) + 
    geom_point(
      data = dplyr::filter(doublet_df, {{pred_column}} == "doublet"), 
      aes(color = {{color_column}})
    )
  }

  return(p)
}

@@

```{r}
dub_tab <- print_doublet_table(doublet_df, "ground_truth", is_ground_truth = TRUE)
dub_tab$table
```

```{r results = "asis"}
dub_tab$percent
```

```{r}
plot_pca(doublet_df, ground_truth, color_lab = "Ground truth doublets")
```

@@

```{r}
dub_tab <- print_doublet_table(doublet_df, "scdbl_prediction")
dub_tab$table
```

```{r results = "asis"}
dub_tab$percent
```

@@

```{r fig.width = 8}

d_plot <- plot_pca(
  doublet_df, 
  color_column = scdbl_prediction,
  color_lab = "scDblFinder prediction"
)

c_plot <- plot_pca(
  doublet_df, 
  color_column = scdbl_score,
  pred_column = scdbl_prediction,
  color_type = "c",
  color_lab = "scDblFinder score"
)

@@

```{r}
dub_tab <- print_doublet_table(doublet_df, "scrublet_prediction")
dub_tab$table
```

```{r results = "asis"}
dub_tab$percent
```

@@

```{r fig.width = 8}
d_plot <- plot_pca(
  doublet_df, 
  color_column = scrublet_prediction,
  color_lab = "scrublet prediction"
)

c_plot <- plot_pca(
  doublet_df, 
  color_column = scrublet_score,
  pred_column = scrublet_prediction,
  color_type = "c",
  color_lab = "scrublet score"
)

@@

      plot_pca(
        doublet_df, 
        color_column = !!as.symbol(col_name),
        color_lab = glue::glue("cxds prediction at {threshold}")
      ) +
      labs(title = glue::glue("cxds prediction at {threshold}"))
    }
)
wrap_plots(prediction_pcas)
```

```{r}
# just need a single score pca for this case
# we'll provide the least stringent predictions; this only affects point layering in the plot
plot_pca(
  doublet_df, 
  color_column = cxds_score, 
  pred_column = cxds_prediction_0.5,
  color_type = "c",
  color_lab = glue::glue("cxds scores")
)
```

@@

```{r}
dub_tab <- print_doublet_table(doublet_df, "cxds_prediction_0.5")
dub_tab$table
```

```{r results = "asis"}
dub_tab$percent
```

@@

```{r}
dub_tab <- print_doublet_table(doublet_df, "cxds_prediction_0.75")
dub_tab$table
```

```{r results = "asis"}
dub_tab$percent
```

@@

```{r}
dub_tab <- print_doublet_table(doublet_df, "cxds_prediction_0.9")
dub_tab$table
```

```{r results = "asis"}
dub_tab$percent
```
